---
title: "Chapter 1"
author: "Laura"
date: "8/7/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---
https://jrnold.github.io/r4ds-exercise-solutions/data-visualisation.html Website w/answers
https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf shortcuts for markdown


library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
  
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

ggplot(data = mpg) +
  geom_point(mapping =aes(x = displ, y = hwy, size = class))

# Top 
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
  
# Bottom 
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
  
ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```{r}
library(ggplot2)
ggplot(data = mpg)
```

```{r}
# I want to see mpg data

head(mpg)
str(mpg)
write.csv(mpg, file = "mpg")

```

I am confused. I am going to try to keep going and see if it starts to make more sense. 

Facets:

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

Did not display at first in plots pane, Googled and found to click on gear button in this pane and choose "Chunk output in console" and it is fixed!

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

Without rows or columns:


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(. ~ cyl)
```

3.5.1 Exercises:
1) What happens if you facet on a continuous variable?

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  facet_grid(. ~ cty)
```
The continuous variable changed to a catagorical variable, and each number got its own facet. (looked at book answers for term "catetorical" variable)

2) What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cty)) +
  facet_grid(drv ~ cyl)
```
Blank plots are where there's no type of that car: so for example no 4 wheel drive 5 cylinder vehicles. 

3) What plots does the following code make? What does . do?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```


It ignored the part repalced with the . in the facet. Answer key says ignored that "dimension"

4) Take the first faceted plot in this section:

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

?facet_wrap
It's funny that I searched this before the question asking me to. 

Advantages to faceting: pulls data into separate, easier to see variables, but could be harder to compare overall picture. If had a large dataset, may be best to have it faceted so there's meaning to the dots as opposed to all squished and potentially overlapping in one graph. 

## Geometric Objects

```{r}
# left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

```
```{r}
# right
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

You can change the line type in line graphs:

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```
That separated the drive types compared to the other where they were combined in one line (I think)

So Many geoms, like geom_smooth(), use a single geometric object to display multiple rows of data


```{r}
geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```

```{r}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = TRUE)
```

Above did not work! 

To display multiple geoms in the same plot, add multiple geom functions to ggplot()

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

Another error message: Error: Cannot add ggproto objects together. Did you forget to add this object to a ggplot object?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```


# Exercises:

2) Predict what this code will look like

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, colour = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

5)Will these two graphs look different? Why/why not?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```
geom_point() and geom_smooth() use the same data and mappings so look the same



```{r}
install.packages("maps")

library(maps)
library(ggplot2)
nz <- map_data("usa")

 

ggplot(nz, aes(long, lat, group = group)) +

  geom_polygon(fill = "white", colour = "black")

 

ggplot(nz, aes(long, lat, group = group)) +

  geom_polygon(fill = "white", colour = "black") +

  coord_quickmap()
```


Didn't work because could not find nz

```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```

Neat! the above worked fine

## Exercises
1) Turn a stacked bar chart into a pie chart using coord_polar()

```{r}
ggplot(mpg, aes(x = factor(1), fill = drv)) +
  geom_bar()
```
```{r}
ggplot(mpg, aes(x = factor(1), fill = drv)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y")
```

```{r}
?labs()
```

Labs modify axes, legends, and plot tables (y, x, title, subtitle, caption, etc.)

3) What's the difference between coord_quickmap() and coord_map()?

```{r}
?coord_quickmap
```

4) what does the following plot tell you about the relationship between city and hwy mpg? why is coord_fixed() important? What does geom_abline() do?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()

```

```{r}
?coord_fixed
```

```{r}
?geom_abline
```





