---
title: "Chapter 3.0"
author: "Laura"
date: "9/3/2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(nycflights13)
library(tidyverse)
```

```{r}
flights
str(flights)
```
```{r}
View(flights)
```
I like how that made the data look :) 


```{r}
Filtered1_1  <- filter(flights, month == 1, day == 1)
```
```{r}
View(Filtered1_1)
```
Wanted to see if View would work, error in view : unused argument for (day == 1) hmmm. 

```{r}
?View
```

To save result, need to use assignment operator <- 

```{r}
jan1 <- filter(flights, month == 1, day == 1)
```

moves to environment w/title "jan1" 

```{r}
(dec25 <- filter(flights, month == 12, day == 25))
```

adding parentheses the variale and prints results if you add () (above)


```{r}
nov_dec <- filter(flights, month %in% c(11, 12))
```


Nothing comes up using the above



```{r}
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

## Exercises: 
Find all flights that

1) Had an arrival delay of two or more hours

(make sure to use arr_delay for this one:)
```{r}
filter(flights, arr_delay >= 120)
```


2) Flew to Houston (IAH or HOU)

```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```


Did not put quotes in at first, didn't work. Not sure how to type "or" line. Will google that now. 
* It's above the "enter" key: shift + \ = | 

3) Were operated by United, American, or Delta) 

```{r}
filter(flights, carrier %in% c("AA", "DL", "UA"))
```

So using %in% and then c("whatever you want", "whatever else you want") brings up same thing as saying == "this" | == "this" | == "this too", etc. 

4) Departed in summer (July, Aug, Sept)

```{r}
filter(flights, month %in% c("7", "8", "9"))
```

Can also use: 

```{r}
filter(flights, month %in% c(7:9))
```



5) flights that arrived more than 2 hrs late but didn't leave late

```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)
```
Using comma shows "and" needed for data to match request (ie needs to be both, no need to say this "AND" this for resutls)

6) Were delayed by at least an hour, but made up over 30 minutes in flight.

```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```

So this says that the dep delay was 60 min or more, and the departure delay minus the arriva delay was greater than 30 min. 


7) Departed between midnight and 6am (inclusive)

```{r}
filter(flights, dep_time <= 600 | dep_time == 2400)
```

Can't just say <= 600 because of the way the time data is written w/midnight being 2400. So saying <=600 | == 2400 gets all times between midnght and 6am included. 

Tried knitting this to see but error in line 59 as I have noted as well. Keeping for now to look at more closely later. 

8) what is the filter helper between()? 

```{r}
?between
```
Between filters numbers between what you set, so instead of saying x>= left & x<= right, you can say between(x, left, right)

For example for the summer months question:

```{r}
filter(flights, between(month, 7, 9))
```

9) How many flights have a missing dep_time? What other variables are missing? What might these rows represent? 

```{r}
filter(flights, is.na(dep_time))
```

Used is.na to figure out (p49 in book) The arrival time is also NA, these are flights that were cancelled.

10) why is NA ^0 not missing? Why is NA | TRUE not missing? WHy is FALSE & NA not missing? can you figure out the general rule?

```{r}
NA ^ 0
```

```{r}
NA | TRUE
```

```{r}
FALSE & NA
```

A bit confused about what these numbers mean. 


# Arrange Rows with arrange()

```{r}
arrange(flights, year, month, day)
```

```{r}
arrange(flights, desc(arr_delay))
```

1) how could you use arrange() to sort all missing values to the start? (hint: use is.na())


```{r}
arrange(flights, desc(is.na(dep_time)), dep_time)
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest. 

```{r}
arrange(flights, desc(dep_delay))
```
These are the flights that were most delayed

```{r}
arrange(flights, dep_delay)
```

w/out descending it shows in ascending order.

3. Sort flights to find the fastest flights.

```{r}
arrange(flights,  )
```
No idea how to do this...


4. Which flight traveled the longest? the shortest? 

```{r}
arrange(flights, desc(distance))
```

So that's for the longest, shortest are:

```{r}
arrange(flights, distance)
```

I laughed out loud at the 17 mi flight

## Select item with select()

select() allows you to rapidly zoom in on a useful subset using operations based on the names of the variables

```{r}
select(flights, year, month, day)
```

```{r}
select(flights, year:day)
```


Ok yep pretty straight forward

```{r}
select(flights, -(year:day))
```


?select()

rename flights using rename() because it keeps all other variables. Using select() would drop all other values not explicitly detailed

```{r}
rename(flights, tail_num = tailnum)
```

OR use select() in conjunction with everything() as follows:

```{r}
select(flights, time_hour, air_time, everything())
```

So it selected everything after the ones I named of time_hour and air_time. 

```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
```

```{r}
select(flights, 4, 6, 7, 9)

```

2) What happens if you include the name of a variable multiple times in a select() call? 

```{r}
select(flights, dep_time, dep_time)
```

It ignores repeats, no error messages occur. Useful b/c you can use select() w/ everything() to change order of columns w/out having to specify the names of all of them

3) What does the one_of() function do? Why might it be helpful in conjunction with this vector?


```{r}
? one_of
```

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```

Selects variables w/a character vector rather than unquoted variable name arguments

4) Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```
it ignores case

```{r}
select(flights, contains("TIME", ignore.case = FALSE))
```

adding ignore.case = FALSE will change it to use case. 


## Add new variables with mutate() 

```{r}
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)
```

So that added the gain & speed columns. 


You can refer to new columns you've created:

```{r}
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

If you only want to keep the new variables, use transmute():

```{r}
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```



#Page 68 in Da book: 


* Mean and Median 

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0]) # the average positive delay
  )
```

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```

Got this error both things used today: Error in eval(lhs, parent, parent) : object 'not_cancelled' not found

Not sure what's up. 

Counts 

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```


Hmm. Ok found part way back that names delays as not_cancelled 

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```

Thought that might name the not cancelled and work. Same error though. 


```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
```

Ok, I had to go back to this one and now it's working. I'm going to go back and see if the others work as well. 

# Grouping by Multiple Variables

```{r}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))
```

```{r}
(per_month <- summarise(per_day, flights = sum(flights)))
```

```{r}
(per_year  <- summarise(per_month, flights = sum(flights)))
```


```{r}
daily %>% 
  ungroup() %>%             # no longer grouped by date
  summarise(flights = n())  # all flights
```


# Grouped Mutates (& filters)

Find the worst members of each group
```{r}
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)
```































